Binary companions of red giant stars I: \\ catalog and companions that survive the common envelope

### 2017-08-14 paper/1-catalog/twoface-catalog.tex
refs and text

@@

\subsection{Detection and orbit fitting}\label{sec:fitting}

Our approach here is to proceed in two phases.
In the first phase, we obtain a posterior sampling in binary-system
parameter space for every individual star, treating it as a
single-lined (SB1) spectroscopic binary system with a single
companion.
This sampling is performed under a relatively uninformative prior pdf,
which will be called the ``\emph{interim prior}''.
These samplings are used to discover and characterize individual binary-star
systems, and generate (among other things) a catalog of binaries.
In the second phase, we perform a hierarchical inference of the binary
population, using (responsibly) these samplings under the interim
prior as inputs.
In the second phase we use all samplings of all stars, not just those that
pass our thresholds for discovery as binary systems.
In this \sectionname, we describe the first phase, that is, the
individual-system sampling phase for discovery and characterization.

We perform our individual system fits---that is, make our posterior
samplings---under the following assumptions:
\begin{description}
\item[no trinaries] whatevs: No trinaries or beyond. Obviously wrong!
\item[SB1] whatevs: There is no significant light contributed by the secondary.
\item[Kepler] whatevs: No non-gravitational contributions to RV history.
\item[noise model] whatevs: Measurements are unbiased and noise estimates are correct up to jitter; everything Gaussian.
\item[interim prior] We adopt a prior pdf on binary-system parameters
  with the following properties...
\end{description}

Summary of \thejoker\ and what it does; that is, that it obeys the
relevant parts of our assumptions.

Any modifications or knob-settings for \thejoker?

>>>>

\todo{An alternate approach to this intro is to start with an overview of
      binary stars, radial velocity measurements of binarity in general...}

\todo{As is, the intro is missing many citations}

@@

\subsection{Orbit fitting and velocity modeling}\label{sec:fitting}

For every source in the sample of \apogee\ stars defined in
\sectionname~\ref{sec:data}, we obtain a posterior sampling in binary-system
parameter space, treating it as a single-lined (SB1) spectroscopic binary system
with a single companion.
This sampling is performed under a relatively uninformative prior pdf, and the
resulting posterior samplings are used to discover and characterize individual
binary-star systems and generate a catalog of companions.
We use \thejoker\ to perform the posterior samplings; we briefly describe the
algorithm below, but a full description is given in previous work
(\citealt{Price-Whelan:2017}).
In this \sectionname, we describe the assumptions and method used to generate
individual-system samplings.

We perform our individual system fits---that is, make our posterior
samplings---under the following assumptions:
\begin{description}
\item[no multiplets] All radial-velocity variations of the primary star are
  induced by a single companion.
  This is motivated by the idea that triple star systems are usually
  hierarchical so that the period of the inner binary is typically much shorter
  than the orbital period of the outer body (\todo{CITATION NEEDED}).
  At present, we ignore the possibility of higher-order multiple systems.
  \todo{revisit this - do we want to also sample over a long-term trend?}
\item[Kepler] Related to the first point, all velocity variations of the primary
  are gravitational, and we therefore ignore the possibility of intrinsic
  variation from, e.g., stellar oscillations.
\item[SB1] All spectra are single-lined; that is, we assume that the secondary
  is significantly fainter and is thus undetected in the spectra.
  This assumption is motivated by the fact that we expect \RGB\ stars to be
  substantially more luminous than their typical companion.
  However, there are known double-lined binary stars in the \apogee\ catalog
  (El-Badry et al., in prep.), and an unknown fraction of \RGB--\RGB\ binaries.
\item[simple noise model] Measurements are unbiased and noise estimates are
  correct up to an \todo{unknown} extra variance (the ``jitter'').
  All noise contributions result in Gaussian uncertainties on the individual
  radial-velocity measurements.
\end{description}

\thejoker\ is a custom-built Monte Carlo sampler designed to produce independent
posterior samples in orbital parameters, given radial-velocity measurements
under the assumptions above.
Our parametrization is similar to the notation in \citet{Murray:2010}:
The radial velocity $v$ at time $t$ is given by
\begin{equation}
  v(t;\bs{\theta}) = v_0 + K\,[\cos\left(\omega + f(t; e, P, \phi_0)\right) +
    e\,\cos\omega]
\end{equation}
where $\bs{\theta} = (P,e,\phi_0,\omega,K,v_0)$---period, eccentricity, initial
phase, argument of pericenter, velocity semi-amplitude, Barycentric velocity---
and the true anomaly, $f$, is a function of the specified parameters (see
\sectionname~2 of \citealt{Price-Whelan:2017} or \eqname~63 in
\citealt{Murray:2010}).

\thejoker\ was designed for the extremely multi-modal pdfs expected when the
number of radial-velocity measurements of a source is small, or the data have
low signal-to-noise.
While other Markov Chain Monte Carlo (MCMC) methods have difficulty producing
independent samples with such data, \thejoker\ succeeds by brute force.
After generating an initial (very large) library of prior samples from an
assumed prior pdf (see below), the (typically multi-modal) likelihood is
evaluated at each sample and used to rejection sample.
In practice, given a number of requested samples for each star, the sampling
proceeds iteratively: since it is easier to accept samples when the data is
sparse or noisy, far more prior sample draws (and thus likelihood
evaluations) must occur under very constraining data.

In addition to generating posterior samples for each source, we also compute the
fully marginalized likelihood (FML) under two other models: (1) a model in which
the radial velocity of the source is constant, and (2) a model in which the
radial-velocity variations are purely linear.
We later use these FML values to define cuts on our full catalog of posterior
samples to produce a catalog of companions.

\todo{equations?}
